The universal property of infinite direct sums in C$^*$-categories and W$^*$-categories

### 2019-09-03 direct_sums.tex
comments reviewer

\thanks{\textit{Acknowledgements.} We thank Robert Furber for copious help and discussion, as well as Chris Heunen and Martti Korvanen for discussion. Part of this work was conducted while the first author was at the Max Planck Institute for Mathematics in the Sciences in Leipzig, Germany.}

@@

Traditionally, universal properties have been formulated mostly for purely algebraic structures only. Treating analytical structures categorically is straightforward in some cases, but very challenging in many other cases. It usually involves equipping the categories under consideration themselves with certain analytical structures. An interesting case of where this happens is the theory of \emph{W$^*$-categories}, which model categories of Hilbert space representations of C$^*$-algebras or W$^*$-algebras. The most paradigmatic example of a W$^*$-category being $\Hilb$, the category of Hilbert spaces with bounded linear maps.

@@

\begin{defn}
    Let~$\cat{C}$ be a category together with an endofunctor~$*\,\colon \cat{C} \to \cat{C}^\op$,
    whose action on arrows we denote by~$a\mapsto a^*$. The pair $(\cat{C},*)$ is called a \emph{complex~$*$-category} if the following conditions hold.
\begin{enumerate}
    \item $\cat{C}$ is enriched over complex vector spaces, meaning that:
	    \begin{enumerate}
		    \item every homset~$\cat{C}(A,B)$ is a complex vector space,
		    \item composition of morphisms is bilinear.
	   \end{enumerate}
    \item The contravariant endofunctor~$*$ is
	    \begin{enumerate}
		    \item identity-on-objects: $a^* : B \to A$ for $a : A \to B$,
		    \item involutive: $a^{**} = a$,
		    \item conjugate linear:
			    \[
				    (\lambda a)^* \ =\  \overline{\lambda} a^*,\qquad (a+b)^* \ =\  a^* + b^*
			    \]
			    for morphisms $a, b : A \to B$ and $\lambda\in\C$.
	    \end{enumerate}
    \item For every morphism~$a : A \to B$,
	    \begin{enumerate}
		    \item there is a $b :  A \to A$ with~$a^*a=b^*b$,
		    \item $a^*a=0$ implies $a = 0$.
	    \end{enumerate}
\end{enumerate}
\end{defn}

@@

\begin{defn}
A complex $*$-category $(\cat{C},*)$ is called a
\begin{enumerate}
	\item \emph{normed $*$-category} if each homset~$\cat{C}(A,B)$ is a normed space such that
	    \[
		    \| ab \| \ \leq\  \|a\| \, \|b\|
	    \]
                for any~$a\colon B \to C$ and~$b\colon A \to B$.
	\item\label{Cstarcat}
    \emph{C$^*$-category} if moreover every homset is norm-complete, and the C$^*$-identity
    \[
	\|a \|^2\  =\  \|a^*a\|
     \]
	holds for every morphism $a$.
\item\label{Wstarcat}
    \emph{W$^*$-category} if moreover
        where every homset has a Banach space predual.

	Equivalently~\cite[Prop.~2.15]{wstarcat}, a C$^*$-category~$\cat{C}$
            is a~W$^*$-category
            if every~$\cat{C}(A,A)$ is a W$^*$-algebra
            and~$\cat{C}(A,B)$ is self dual
            as a Hilbert~$\cat{C}(A,A)$-module
            with inner product~$\langle a,b\rangle \equiv a^*b$.
\end{enumerate}
\end{defn}

Clearly if we have \ref{Cstarcat}, then every $\cat{C}(A,A)$ is a C$^*$-algebra. Similarly, \ref{Wstarcat} makes $\cat{C}(A,A)$ into a W$^*$-algebra.

@@

\begin{lem}
Let~$a \colon A \leftrightarrows B \colon b$ be any two morphisms
    such that~$\|a \| \leq 1$, $\| b\| \leq 1$ and~$ab=1$.
    Then~$b$ is an isometry with~$a=b^*$.
\label{isometrylemma}
\end{lem}
\begin{proof}
    To start, note~$\|a ^*a\| = \|a\|^2 \leq 1$, so~$a^*a \leq 1$.
    Similarly~$b^*b \leq 1$.
    Combined~$1 = b^*a^*ab \leq b^*b \leq 1$, hence~$b^*b=1$.
    Similarly~$aa^*=1$.
    Clearly~$bb^*bb^* = b 1 b^* = bb^*$, so~$b$ is an isometry.
    Now~$a (1-bb^*) a^* = aa^* - abb^*a^* = 1-1=0$.
    So~$a^* (1-0) a \leq bb^*$ by Lemma~\ref{contrapositionlemma}.
    Similarly~$bb^* \leq a^*a$. Hence~$bb^*=a^*a$.
    We are done: $a = aa^*a=abb^*=b^*$.
\end{proof}

@@

We again let $\cat{C}$ be a C$^*$-category. Let $I$ be a finite or infinite index set. For $(A_i)_{i\in I}$ a family of objects in $\cat{C}$, we consider the functor
\[
	\bigoplus_i \cat{C}(A_i,-) \colon \  \cat{C} \to \Ban
\]
given by associating to every object $B$ the set
\beq
\label{AiB}
	\bigoplus_i \cat{C}(A_i , B) \ := \ \Bigl\{ (f_i : A_i \to B)_{i\in I} \Bigm\vert\textstyle  \sum_i f_i f_i^* < \infty \Bigr\}.
\eeq
Here, $\sum_i f_i f_i^*$ can be interpreted as converging in the ultraweak topology\footnote{For positive elements~$(x_i)_{i \in I}$ of a W$^*$-algebra, convergence and value of the sum~$\sum_{i \in I} x_i$ coincides in the ultraweak and ultrastrong topology and, in fact, equals the supremum of the partial sums.} in the case where $\cat{C}(B,B)$ is a W$^*$-algebra. If this is not the case, what we mean by the condition $\sum_i f_i f_i^* < \infty$ is that there is a fixed element which upper bounds all partial sums $\sum_{i\in S} f_i f_i^*$ for finite $S\subseteq I$, or equivalently that the sum converges ultraweakly in the double dual W$^*$-algebra $\cat{C}(B,B)^{**}$.

The set~\eqref{AiB} is a vector space under componentwise addition and scalar multiplication, where closure under addition follows from $f_i g^*_i + g_i f^*_i \leq f_i f^*_i + g_i g^*_i$, which holds because of $(f_i - g_i) (f^*_i - g^*_i) \geq 0$. We furthermore consider this vector space as a normed space under the norm
\[
	\| (f_i)_{i \in I} \| \ :=\  \Bigl\| \sum_i f_i f_i^* \Bigr\|^{1/2},
\]
where we interpret the sum as before either in terms of a supremum over the norm of finite partial sums, or as converging in the double dual $\cat{C}(B,B)^{**}$.
The triangle inequality for this norm is not directly obvious.
t follows from the following generalization of the
    Cauchy--Schwarz inequality, see e.g.~\cite[\S142\textsubscript{III \& IV}]{bas},
\begin{equation*}
    \Bigl(\sum_i f_i g_i^*\Bigr) \Bigl(\sum_j g_j f_j^*\Bigr)
    \ \leq \ \Bigl\| \sum_i g_ig_i^* \Bigr\| \: \sum_j f_j f_j^* .
\end{equation*}

@@

\begin{proof}
Suppose that we are given a Cauchy sequence of families $(f^{(n)})_{n\in\N} = ( ( f_i^{(n)} )_{i\in I} )_{n\in\N}$. For given $j$, the projection map
\[
	\bigoplus_i \cat{C}(A_i,B) \longrightarrow \cat{C}(A_j,B)	
\]
is bounded, and therefore $( f_j^{(n)} )_{n\in\N}$ is Cauchy as well; let us denote the limit by $f_j : A_j \to B$. We will now show that $f \in \bigoplus_i \cat{C}(A_i,B)$ for $f := (f_j)_{j\in I}$, which amounts to verifying the relevant square summability. The Cauchy assumption implies that the sequence $(f^{(n)})_{n\in\N}$ is uniformly bounded by some constant $C$,
\[
    \| f^{(n)} \| \ = \ 
	\Bigl\| \textstyle \sum_i f_i^{(n)} (f_i^{(n)})^* \Bigr\|^{1/2} \ \leq\  C.
\]
Restricting to a finite partial sum over $S\subseteq I$ and taking $n\to\infty$ shows that
\[
	\Bigl\| \textstyle \sum_{i\in S} f_i f_i^* \Bigr\|^{1/2} \ \leq\  C
\]
as well. Since $S$ was arbitrary, we conclude that $\| f \| \leq C$.

The convergence $f_i \to f$ works similarly. By assumption, for every~$\eps > 0$ we have~$\left\| f^{(n)} - f^{(m)} \right\| \leq \eps$ for almost all $n,m\in\N$. In particular, we know that for any finite $S \subseteq I$,
\[
	\Bigl\| \sum_{i\in S} \bigl(f_i^{(n)} - f_i^{(m)}\bigr) \bigl(f_i^{(n)} - f_i^{(m)}\bigr)^* \Bigr\|^{1/2} \ \leq\  \eps
\]
for almost all $n,m\in\N$. Taking $m\to\infty$ gives
\[
	\Bigl\| \sum_{i\in S} \bigl(f_i^{(n)} - f_i\bigr) \bigl(f_i^{(n)} - f_i\bigr)^* \Bigr\|^{1/2} \ \leq \ \eps.
\]
Since $S$ was arbitrary, we conclude $\| f^{(n)} - f \|  \leq \eps$ for almost all $n$.
\end{proof}

@@

\begin{thm}
\label{directsum_equiv}
Let $\cat{C}$ be a W$^*$-category and $(A_i)_{i\in I}$ a family of objects in $\cat{C}$. Then the following pieces of structure are equivalent for an object $A$:
\begin{enumerate}
\item\label{universalprop} $A$ is an $I$-indexed direct sum $\bigoplus_i A_i$.
\item\label{universalprop2} 
There is a family of
        morphisms~$\left(\kappa_j : A_j \to A\right)_{j\in I}$
        with~$\sum_{j \in I} \kappa_j \kappa_j^* < \infty$
        such that for any object~$B$
        and family of
        morphisms~$\left(f_j: A_j \to B\right)_{j\in I}$
        with~$\sum_{j \in I}f_j f_j^* < \infty$,
        there is a unique~$f\colon A \to B$
        such that $f \kappa_j = f_j$ for all~$j \in I$, and $\| f \|^2 = \| \sum_{j\in I} f_j f_j^* \|$.
\item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to A\right)_{j\in I}$ such that $\kappa_i^*\kappa_j = \delta_{ij}$, and
\beq
\label{complete}
        \sum_{j\in I} \kappa_j \kappa_j^* \ =\  1_A.
\eeq
\end{enumerate}
\end{thm}

@@

We finally show that \ref{biprodsum} implies \ref{universalprop2}. So let~$\left(f_j: A_j \to B\right)_{j\in I}$ be any family of morphisms
        with~$\sum_{j \in I}f_jf_j^* < \infty$.
The homset~$C(B,A)$
    is a self-dual Hilbert~$\cat{C}(B,B)$-module
    with~$\cat{C}(B,B)$-valued inner
    product~$\langle g, f\rangle \equiv g^*f$~\cite[Proposition~2.15]{wstarcat}.
The self-duality is equivalent to the fact
    that~$\cat{C}(B,A)$ is complete in the \emph{ultranorm topology};
    this is the topology
    generated by the seminorms~$\| f \|_\omega \equiv \omega(\langle f,f\rangle)^\frac{1}{2}$
    indexed by normal states~$\omega$ on~$\cat{C}(B,B)$~\cite[\S149\textsubscript{V}]{bas}, also known as the \emph{s-topology}~\cite[p.~88]{wstarcat}. We now turn to the construction of $f : A \to B$.
    Pick any normal state~$\omega$ on~$\cat{C}(B,B)$
        and finite subset~$S \subseteq I$.  We have
\begin{alignat*}{2}
    \bigl\| \sum_{j \in S} \kappa_j f_j^* \bigr\|^2_\omega
    &\ \equiv \ 
    \omega\Bigl(\bigl(\sum_{i \in S} \kappa_i f_i^*\bigr)^*
            \bigl(\sum_{j \in S} \kappa_j f_j^*\bigr)\Bigr) \\
    &\ =\ 
    \omega\Bigl( \sum_{j \in S} f_j f_j^*
    \Bigr) &\qquad&\text{as~$\kappa_j^*\kappa_i = \delta_{ij}$}.
\end{alignat*}
From this and~$\sum_{j \in I} f_jf_j^* < \infty$,
    it follows that the net~$\left(\sum_{j \in S} \kappa_j f_j^*\right)_S$
    is ultranorm Cauchy as $S$ ranges over all finite subsets of $I$.
    Define~$f := \left( \sum_{j \in I} \kappa_jf_j^* \right)^*$,
        where the sum is understood to converge ultranorm.
The assignment~$g \mapsto \kappa_j^*g$
    yields a bounded~$\cat{C}(B,B)$-linear map between
    the Hilbert~$\cat{C}(B,B)$-modules~$\cat{C}(B,A)$ and~$\cat{C}(B,A_j)$
    and is therefore ultranorm continuous~\cite[\S148]{bas}.
    Hence
        $\kappa_j^* f^*
             =  \kappa_j^* \sum_{i} \kappa_i f_i^*
             =  \sum_i \kappa_{j }^*\kappa_i f_i^*
            =  f_j^*$
            and so~$f \kappa_j = f_j$ as desired.
Next we will show that~$\|f \|^2 = \|\sum_j f_j f_j^*\|$.
    By definition of~$f^*$ and \cite[\S148\textsubscript{V}]{bas}
         we have~
 \begin{equation*}
         \langle f^*, f^*\rangle
         \ =\  \uwlim_{S\subseteq I\text{ finite}} \, \Bigl\langle \sum_{j \in S} \kappa_j f^*_j,
                        \sum_{j \in S}\kappa_j f^*_j \Bigr\rangle
         \ =\  \uwlim_{S\subseteq I\text{ finite}} \, \sum_{j \in S} f_j f_j^*
        \ = \ \sum_j f_j f_j^*
 \end{equation*}
and so~$\| f \|^2 = \|f^*\|^2 = \|\langle f,f \rangle \|
                = \|\sum_j f_j f_j^* \|$ as desired.
Finally, to show uniqueness,
    assume~$f'\colon A \to B$ is any (other)
    morphism with~$f' \kappa_i = f_i$.
    Note that~$\sum_j \kappa_j \kappa_j^*$
       converges ultrastrongly
       and so ultranorm as well.
With similar reasoning as before,
        the composition map~$g \mapsto f' g$ is ultranorm continuous,
        hence
\begin{equation*}
    f'
    \ =\ f' \sum_{j \in I} \kappa_j \kappa_j^*
    \ =\ \sum_{j \in I} f' \kappa_j \kappa_j^*
    \ =\ \sum_{j \in I} f_j \kappa_j^*
    \ =\ \sum_{j \in I} f \kappa_j \kappa_j^*
    \ =\ f,
\end{equation*}
    where the sums converge ultranorm.
\end{proof}

>>>>

\thanks{\textit{Acknowledgements.} We thank Robert Furber for copious
help and discussion, as well as Chris Heunen and Martti Karvonen
for discussion. Part of this work was conducted while the first
author was at the Max Planck Institute for Mathematics in the
Sciences in Leipzig, Germany.}

@@

There appear many useful universal properties for purely
algebraic structures.  Treating analytical structures
categorically is just as straightforward in some cases, but very challenging
in many other cases. It usually involves equipping the categories
under consideration themselves with certain analytical structures.
An interesting case of where this happens is the theory of
\emph{W$^*$-categories}, which model categories of Hilbert space
representations of C$^*$-algebras or W$^*$-algebras. The most
paradigmatic example of a W$^*$-category being $\Hilb$, the category
of Hilbert spaces with bounded linear maps.

@@

\begin{defn}
    Let~$\cat{C}$ be a category together with an endofunctor~$*\,\colon \cat{C} \to \cat{C}^\op$,
    whose action on arrows we denote by~$a\mapsto a^*$. The pair $(\cat{C},*)$ is called a \emph{complex~$*$-category} if the following conditions hold.
\begin{enumerate}
    \item $\cat{C}$ is enriched over complex vector spaces, meaning that:
	    \begin{enumerate}
		    \item every homset~$\cat{C}(A,B)$ is a complex vector space,
		    \item composition of morphisms is bilinear.
	   \end{enumerate}
    \item The contravariant endofunctor~$*$ is
	    \begin{enumerate}
		    \item identity-on-objects: $a^* : B \to A$ for $a : A \to B$,
		    \item involutive: $a^{**} = a$,
		    \item conjugate linear:
			    \[
				    (\lambda a)^* \ =\  \overline{\lambda} a^*,\qquad (a+b)^* \ =\  a^* + b^*
			    \]
			    for morphisms $a, b : A \to B$ and $\lambda\in\C$.
	    \end{enumerate}
    \item For every morphism~$a : A \to B$,
	    \begin{enumerate}
		    \item there is a morphism $b :  A \to A$ with~$a^*a=b^*b$,
		    \item $a^*a=0$ implies $a = 0$.
	    \end{enumerate}
\end{enumerate}
\end{defn}

@@

\begin{defn}
A complex $*$-category $(\cat{C},*)$ is called a
\begin{enumerate}
	\item \emph{normed $*$-category} if each homset~$\cat{C}(A,B)$ is a normed space such that
	    \[
		    \| ab \| \ \leq\  \|a\| \, \|b\|
	    \]
                for any~$a\colon B \to C$ and~$b\colon A \to B$.
	\item\label{Cstarcat}
    \emph{C$^*$-category} if moreover every homset is norm-complete, and the C$^*$-identity
    \[
	\|a \|^2\  =\  \|a^*a\|
     \]
	holds for every morphism $a$.
        (This turns $\cat{C}(A,A)$ into a~C$^*$-algebra.)
\item\label{Wstarcat}
    \emph{W$^*$-category} if moreover
        where every homset has a Banach space predual.

	Equivalently~\cite[Prop.~2.15]{wstarcat}, a C$^*$-category~$\cat{C}$
            is a~W$^*$-category
            if every~$\cat{C}(A,A)$ is a W$^*$-algebra
            and~$\cat{C}(A,B)$ is self dual
            as a Hilbert~$\cat{C}(A,A)$-module
            with inner product~$\langle a,b\rangle \equiv a^*b$.
\end{enumerate}
As the homset~$\cat{C}(A,A)$
    in a C$^*$-category
    is a C$^*$-algebra,
    we will freely use definitions from~C$^*$-algebras on endomorphisms,
    so we say~$a\colon A \to A$ is a \emph{projection}
        if it is a projection as an element of the~C$^*$-algebra
        $\cat{C}(A,A)$.
Several notions extend naturally to all morphisms:
    a morphism~$a \colon A \to B$ is called an \emph{isometry} if~$a^*a=1$
        (and in that case~$aa^*$ is a projection)
        and it is a \emph{unitary} if additionally~$aa^*=1$.
\end{defn}

@@

\begin{lem}
Let~$a \colon A \leftrightarrows B \colon b$ be any two morphisms
    such that~$\|a \| \leq 1$, $\| b\| \leq 1$ and~$ab=1$.
    Then~$b$ is an isometry with~$a=b^*$.
\label{isometrylemma}
\end{lem}
\begin{proof}
    To start, note~$\|a ^*a\| = \|a\|^2 \leq 1$, so~$a^*a \leq 1$.
    Similarly~$b^*b \leq 1$.
    Combined~$1 = b^*a^*ab \leq b^*b \leq 1$, hence~$b^*b=1$
        and so~$b$ is an isometry.
    Similarly~$aa^*=1$.
    Now~$a (1-bb^*) a^* = aa^* - abb^*a^* = 1-1=0 \leq 0$.
    So~$a^* (1-0) a \leq bb^*$ by Lemma~\ref{contrapositionlemma}.
    Similarly~$bb^* \leq a^*a$. Hence~$bb^*=a^*a$.
    We are done: $a = aa^*a=abb^*=b^*$.
\end{proof}

@@

We again let $\cat{C}$ be a C$^*$-category. Let $I$ be a finite or infinite index set. For a family $(A_i)_{i\in I}$ of objects in $\cat{C}$, we consider the functor
\[
	\bigoplus_i \cat{C}(A_i,-) \colon \  \cat{C} \to \Ban
\]
given by associating to every object $B$ the set
\beq
\label{AiB}
	\bigoplus_i \cat{C}(A_i , B) \ := \ \Bigl\{ (f_i : A_i \to B)_{i\in I} \Bigm\vert\textstyle  \sum_i f_i f_i^* < \infty \Bigr\}.
\eeq
Here, $\sum_i f_i f_i^*$ can be interpreted as converging in the ultraweak topology\footnote{For positive elements~$(x_i)_{i \in I}$ of a W$^*$-algebra, convergence and value of the sum~$\sum_{i \in I} x_i$ coincides in the ultraweak and ultrastrong topology and, in fact, equals the supremum of the partial sums.} in the case where $\cat{C}(B,B)$ is a W$^*$-algebra. If this is not the case, what we mean by the condition $\sum_i f_i f_i^* < \infty$ is that there is a fixed element which upper bounds all partial sums $\sum_{i\in S} f_i f_i^*$ for finite $S\subseteq I$, or equivalently that the sum converges ultraweakly in the double dual W$^*$-algebra $\cat{C}(B,B)^{**}$.

The set~\eqref{AiB} is a vector space under componentwise addition and scalar multiplication, where closure under addition follows from $f_i g^*_i + g_i f^*_i \leq f_i f^*_i + g_i g^*_i$, which holds because of $(f_i - g_i) (f^*_i - g^*_i) \geq 0$. We furthermore consider this vector space as a normed space under the norm
\[
	\| (f_i)_{i \in I} \| \ :=\  \Bigl\| \sum_i f_i f_i^* \Bigr\|^{1/2},
\]
where we interpret the sum as before either in terms of a supremum over the norm of finite partial sums, or as converging in the double dual $\cat{C}(B,B)^{**}$.
(This norm is in fact the one associated to the~$B$-valued inner
        product~$\langle (f_i)_i, (g_i)_i \rangle  :=
                \sum_i f_ig_i^*$,
                i.e.~$\| (f_i)_i \|^2 = \|\langle (f_i)_i, (f_i)_i \rangle\|$.
        The proof of the triangle inequality
                is similar to that for a
            norm associated to a regular inner product,
            using a generalized Cauchy--Schwarz inequality.
            See for instance~\cite[\S142\textsubscript{III--V}]{bas}
                or~\cite[Prop.~2.3]{paschke}.)

@@

\begin{proof}
Suppose that we are given a Cauchy sequence of families $(f^{(n)})_{n\in\N} = ( ( f_i^{(n)} )_{i\in I} )_{n\in\N}$. For given $j$, the projection map
\[
	\bigoplus_i \cat{C}(A_i,B) \longrightarrow \cat{C}(A_j,B)	
\]
is bounded, and therefore $( f_j^{(n)} )_{n\in\N}$ is Cauchy as well; let us denote the limit by $f_j : A_j \to B$. We will now show that $f \in \bigoplus_i \cat{C}(A_i,B)$ for $f := (f_j)_{j\in I}$, which amounts to verifying the relevant square summability. The Cauchy assumption implies that the sequence $(f^{(n)})_{n\in\N}$ is uniformly bounded by some constant $C$, i.e.
\[
    \| f^{(n)} \| \ = \ 
	\Bigl\| \textstyle \sum_i f_i^{(n)} (f_i^{(n)})^* \Bigr\|^{1/2} \ \leq\  C.
\]
Restricting to a finite partial sum over $S\subseteq I$ and taking $n\to\infty$ shows that
\[
	\Bigl\| \textstyle \sum_{i\in S} f_i f_i^* \Bigr\|^{1/2} \ \leq\  C
\]
as well. Since $S$ was arbitrary, we conclude that $\| f \| \leq C$.

    The convergence $f^{(n)} \to f$ works similarly. By assumption, for every~$\eps > 0$ we have~$\left\| f^{(n)} - f^{(m)} \right\| \leq \eps$ for almost all $n,m\in\N$. In particular, we know that for any finite $S \subseteq I$,
\[
	\Bigl\| \sum_{i\in S} \bigl(f_i^{(n)} - f_i^{(m)}\bigr) \bigl(f_i^{(n)} - f_i^{(m)}\bigr)^* \Bigr\|^{1/2} \ \leq\  \eps
\]
for almost all $n,m\in\N$. Taking $m\to\infty$ gives
\[
	\Bigl\| \sum_{i\in S} \bigl(f_i^{(n)} - f_i\bigr) \bigl(f_i^{(n)} - f_i\bigr)^* \Bigr\|^{1/2} \ \leq \ \eps.
\]
Since $S$ was arbitrary, we conclude $\| f^{(n)} - f \|  \leq \eps$ for almost all $n$.
\end{proof}

@@

\begin{thm}
\label{directsum_equiv}
Let $\cat{C}$ be a W$^*$-category and $(A_i)_{i\in I}$ a family of objects in $\cat{C}$. Then the following pieces of structure are equivalent for an object $A$:
\begin{enumerate}
\item\label{universalprop} $A$ is an $I$-indexed direct sum $\bigoplus_i A_i$.
\item\label{universalprop2} 
There is a family of
        morphisms~$\left(\kappa_j : A_j \to A\right)_{j\in I}$
        with~$\sum_{j \in I} \kappa_j \kappa_j^* < \infty$
        such that for any object~$B$
        and family of
        morphisms~$\left(f_j: A_j \to B\right)_{j\in I}$
        with~$\sum_{j \in I}f_j f_j^* < \infty$,
        there is a unique~$f\colon A \to B$
        such that $f \kappa_j = f_j$ for all~$j \in I$.
Moreover, such an~$f$ satisfies~$\| f \|^2 = \| \sum_{j\in I} f_j f_j^* \|$.
\item\label{biprodsum} There is a family of morphisms $\left(\kappa_j : A_j \to A\right)_{j\in I}$ such that $\kappa_i^*\kappa_j = \delta_{ij}$, and
\beq
\label{complete}
        \sum_{j\in I} \kappa_j \kappa_j^* \ =\  1_A.
\eeq
\end{enumerate}
\end{thm}

@@

We finally show that \ref{biprodsum} implies \ref{universalprop2}. So let~$\left(f_j: A_j \to B\right)_{j\in I}$ be any family of morphisms
        with~$\sum_{j \in I}f_jf_j^* < \infty$.
The homset~$C(B,A)$
    is a self-dual Hilbert~$\cat{C}(B,B)$-module
    with~$\cat{C}(B,B)$-valued inner
    product~$\langle g, f\rangle \equiv g^*f$~\cite[Proposition~2.15]{wstarcat}.
The self duality is equivalent to the fact
    that~$\cat{C}(B,A)$ is complete in the \emph{ultranorm topology};
    this is the topology
    generated by the seminorms~$\| f \|_\omega \equiv \omega(\langle f,f\rangle)^\frac{1}{2}$
    indexed by normal states~$\omega$ on~$\cat{C}(B,B)$~\cite[\S149\textsubscript{V}]{bas}, also known as the \emph{s-topology}~\cite[p.~88]{wstarcat}. We now turn to the construction of $f : A \to B$.
    Pick any normal state~$\omega$ on~$\cat{C}(B,B)$
        and finite subset~$S \subseteq I$.  We have
\begin{alignat*}{2}
    \bigl\| \sum_{j \in S} \kappa_j f_j^* \bigr\|^2_\omega
    &\ \equiv \ 
    \omega\Bigl(\bigl(\sum_{i \in S} \kappa_i f_i^*\bigr)^*
            \bigl(\sum_{j \in S} \kappa_j f_j^*\bigr)\Bigr) \\
    &\ =\ 
    \omega\Bigl( \sum_{j \in S} f_j f_j^*
    \Bigr) &\qquad&\text{as~$\kappa_j^*\kappa_i = \delta_{ij}$}.
\end{alignat*}
From this and~$\sum_{j \in I} f_jf_j^* < \infty$,
    it follows that the net~$(\sum_{j \in S} \kappa_j f_j^*)_S$
    is ultranorm Cauchy as $S$ ranges over all finite subsets of $I$.
    Define~$f := ( \sum_{j \in I} \kappa_jf_j^* )^*$,
        where the sum is understood to converge in the ultranorm topology.
The assignment~$g \mapsto \kappa_j^*g$
    yields a bounded~$\cat{C}(B,B)$-linear map between
    the Hilbert~$\cat{C}(B,B)$-modules~$\cat{C}(B,A)$ and~$\cat{C}(B,A_j)$
    and is therefore ultranorm continuous~\cite[\S148]{bas}.
    Hence
        $\kappa_j^* f^*
             =  \kappa_j^* \sum_{i} \kappa_i f_i^*
             =  \sum_i \kappa_{j }^*\kappa_i f_i^*
            =  f_j^*$
            and so~$f \kappa_j = f_j$ as desired.
Next we will show that~$\|f \|^2 = \|\sum_j f_j f_j^*\|$.
    By definition of~$f^*$ and \cite[\S148\textsubscript{V}]{bas}
         we have~
 \begin{equation*}
         \langle f^*, f^*\rangle
         \ =\  \uwlim_{S\subseteq I\text{ finite}} \, \Bigl\langle \sum_{j \in S} \kappa_j f^*_j,
                        \sum_{j \in S}\kappa_j f^*_j \Bigr\rangle
         \ =\  \uwlim_{S\subseteq I\text{ finite}} \, \sum_{j \in S} f_j f_j^*
        \ = \ \sum_j f_j f_j^*
 \end{equation*}
and so~$\| f \|^2 = \|f^*\|^2 = \|\langle f,f \rangle \|
                = \|\sum_j f_j f_j^* \|$ as desired.
Finally, to show uniqueness,
    assume~$f'\colon A \to B$ is any (other)
    morphism with~$f' \kappa_i = f_i$.
    Note that~$\sum_j \kappa_j \kappa_j^*$
       converges ultrastrongly
       and so in the ultranorm topology as well.
With similar reasoning as before,
        the composition map~$g \mapsto f' g$ is ultranorm continuous,
        hence
\begin{equation*}
    f'
    \ =\ f' \sum_{j \in I} \kappa_j \kappa_j^*
    \ =\ \sum_{j \in I} f' \kappa_j \kappa_j^*
    \ =\ \sum_{j \in I} f_j \kappa_j^*
    \ =\ \sum_{j \in I} f \kappa_j \kappa_j^*
    \ =\ f,
\end{equation*}
    where the sums converge in the ultranorm topology.
\end{proof}
